---
title: "Customer Mindset Metrics, Chapter 6"
subtitle: "Case Study: Zanten Shampoo"
author: "Texbook Author, modified by Jae Min Jung"
date: today
format: 
  html:
    theme: cerulean
    toc: true
    toc-depth: 3
    toc-expand: 1
    toc-location: right-body
    number-sections: true
    code-overflow: wrap
    code-block-border-left: true
    code-line-numbers: true
    embed-resources: true
    df-print: paged
execute:
  echo: fenced
  freeze: true
  error: false
  #cache: true
bibliography: references.bib
---

# Introduction

## Case Background

Zanten is a shampoo brand that has been in the market for a while. Watching the dark clouds over the city of London from her office, Bo Zhang, the Chief Marketing Officer of Sahmpoo brand Zanten, had many thoughts about what to day at tomorrow's six-monthly meeting regarding the somewhat radical changes she was planning to make with her next advertising and promotional decisions.

The next day, Zhang and the Chief Executive Office, Anne Comstock, met over Zoom to discuss the most recent performance outcomes and decide the best possible moves for the future. Zhang began to talk excitedly:

"I am very proud of my team. They have run very successful marketing campaigns over the past few months on social media, giving us a strong position in consumers' hearts. See how we outperform our closest competitor in terms of brand liking." Sharing her screen, she showed Anne Figure 6.1, which shows the trend of liking for Zanten and its closest competitor over the past 24 months.

Zhang continued with a quiet voice: "However, I must admit that the gains in brand liking have not yet been translated into higher revenues. But I believe that these gains in consumer heart share will soon result in better financial performance. We just need to promote ourselves more heavily or increase our advertising substantially, and also believe in ourselves."

The CEO thanked Zhang and said "These analytic insights are great. But there is another pressing issue that we are facing now. The fierce competition is putting a lot of pressure on our margins. As a result, we may need to be a bit careful about how we spend our budget. Therefore, it is extremely important that you justify how these positive changes in consumers' attitudes will turn into better sales performance."

After the meeting, Zhang continued her deliberations on how to allocate marketing resources that would provide noticeable and long-lasting improvements not only in customers' attitudes but also in the brand's sales performance. With one eye on her marketing dashboard and the other on her laptop, she called her analytics team.

"I would like you to prepare a report that provides some concrete directions on how the effectiveness of advertising and promotional campaigns can be improved by examining customer mindset metrics," she said. "I wold like to know (1) to what extent the advertising and promotional campaigns run in recent years have affected consumers' attitudes towards the brand and (2) whether the changes in consumers' perceptions have yielded any significant improvements in our brand's sales performance."

To help Zhang, the marketing analytics team gathered data from various sources.

Moving forward, Zhang has to decide on her promotion and advertising spending choices, and ponders two options:

1.  Quintupling advertising spending from *£100K* to *£500K* or

2.  Doubling promotion spending from *£100K* to *£200K*. S

Which of these options should Zhang propose to the CEO? How would these decisions influence the hearts and minds of consumers, and the sales performance of the company?

## Learning Objectives

These R notes build on the conceptual knowledge on customer attitude metrics. By the end of the application, you will be able to:

1.  Make evidence-based marketing mix decisions by examining the diagnostic value that customer mindset metrics provide.

2.  Quantify the short- and long-run gains in sales that originate from improvement in mindset metrics.

## Setting up the environment

Before we start, let us do some preparation work to get ready.

We need to:

1.  Create a folder (e.g. mindset) under the course directory.

2.  Clone GitHub repository of this chapter at <https://github.com/jaejungca/customer-mindset-metrics> to your local machine under the folder you just created.

3.  The data set for this application, *mindset_shampoo.xlsx* has been moved to "data" folder.

## Packages

```{r include=FALSE}
#| label: load-packages
#Before we load the dataset, we need to install and load a few packages.

if (!require(knitr)) {
  install.packages("knitr", repos = "http://cran.us.r-project.org")
}
if (!require(readxl)) {
  install.packages("readxl", repos = "http://cran.us.r-project.org")
}
if (!require(tseries)) {
  install.packages("tseries", repos = "http://cran.us.r-project.org")
}
if (!require(vars)) {
  install.packages("vars", repos = "http://cran.us.r-project.org")
}
if (!require(R6)) {
  install.packages("R6", repos = "http://cran.us.r-project.org")
}
if (!require(tidyverse)) {
  install.packages("tidyverse", repos = "http://cran.us.r-project.org")
}

library(knitr)
library(readxl)
library(tseries) # for performing time series analysis
library(vars) # for estimating vector autoregressive (VAR) models
library(R6) # for creating classes and objects in R
library(tidyverse)
```

## Data

### Loading the dataset

Next, we will load the dataset to R.

```{r warning=FALSE}
#| label: load-data
# load the dataset
data_shampoo <- read.csv("data/mindset_shampoo.csv")
data_shampoo |> tibble()
```

### Dataset description

We can take a preview of the dataset.

```{r warning=FALSE}
#| label: preview-data
# Take a look at the data

class(data_shampoo)
data_shampoo
```

### Variables

-   **Advertising**: Monthly total spending on advertising media; source: internal company data.

-   **Promotion**: Monthly total spending on promotion; source: internal company data.

-   **Price**: Monthly average price of the shampoo; source: internal company data.

-   **Aided Awareness(% aware)**: % aided awareness is the percentage of respondents who responded Yes to the question "Which of the following brands have you heard of? (Yes/No)"

    -   Online monthly consumer survey conducted by a market research firm.

    -   Respondents are given a list of brands and responds Yes or No to each brand.

    -   The percentage of respondents who said Yes to Zanten is the aided awareness.

-   **Brand Consideration(% considering buying)**: % brand consideration is the percentage of respondents who responded Yes to the question "Which of these brands would you consider buying? (Yes/No)"

    -   Online monthly consumer survey conducted by a market research firm.

    -   Respondents are given a list of brands and responds Yes or No to each brand.

    -   The percentage of respondents who said Yes to Zanten is the consideration.

-   **Brand Liking (average liking)**: Brand liking is measured on a seven-point scale, where 1 means "I don't like it at all" and 7 means "I like it a lot" to a question, "Please indicate how much you like brand X"

    -   The average score of all respondents is the brand liking for that month.

-   **Sales (volume)**: Monthly sales volume of the shampoo (in milliliters); source: internal company data.

::: callout-note
### Plan of analysis

Next, we will start by calculating the *potential*, *stickiness*, and *responsiveness* of these mindset metrics. Then, we will calculate how these metrics contribute to sales (i.e., *conversion*). Finally, by aggregating the results of all these stages we will evaluate the *appeal* of each mindset metric.
:::

# Four criteria

## A. Potential

For awareness and consideration, we need to calculate the average level of awareness and consideration for this shampoo brand and then calculate the potential by deducting one from them.

For liking, things are slightly more complicated since those values are measured on a seven-point scale. Therefore, we need to first convert those values into percentages (e.g. if liking is 3.5 out of 7, then it is actually 50 per cent of the full score and the value should be 50) and then take the average and calculate the corresponding potential.

Let’s reach the following using R:

```{r}
#| label: potential-mindset-metrics
# Calculate potential for awareness and consideration

potential_awareness <- (100 - mean(data_shampoo$awareness)) / 100
potential_consideration <- (100 - mean(data_shampoo$consideration)) / 100

# Calculate potential for liking

data_shampoo$liking_transformed <- data_shampoo$liking / 7
potential_liking <- 1 - mean(data_shampoo$liking_transformed)
```

```{r echo=FALSE}
#| label: potential-table
table_potential <- data.frame(
  potential = c("awareness", "consideration", "liking"),
  value = c(potential_awareness, potential_consideration, potential_liking)
)

table_potential
```

We get that potential is 0.725 for awareness, 0.825 for consideration and 0.263 for liking.

::: callout-note
### **Question 1**:

Judging from what you have now, which mindset metric should you put more effort into? Why?
:::

------------------------------------------------------------------------

## B. Stickiness

*Stickiness*, or inertia, is the **degree to which a change in the level of a metric is upheld over time, without any new stimuli**.

-   This can be modeled by a simple univariate `AR(p)` process on the attitude metric, where stickiness is quantified as *the sum of the AR coefficients*.

-   To do this, we need to run the following code in R.

-   It is important to note that the order of lag for these AR models is determined by *model fit statistics* such as `AIC` or `BIC`.

```{r warning=FALSE}
#| label: stickiness-mindset-metrics
#awareness
ar1 <- ar(data_shampoo$awareness, aic = TRUE)
ar1
summary(ar1)

#consideration
ar2 <- ar(data_shampoo$consideration, aic = TRUE)
ar2

#liking
ar3 <- ar(data_shampoo$liking, aic = TRUE)
ar3
```

From the above results, we can calculate the stickiness of mindset metrics by summing up their AR coefficients for each mindset metric.

```{r}
#| label: stickiness-calculation
stick_awareness <- ar1$ar[1] + ar1$ar[2] + ar1$ar[3]
stick_consideration <- ar2$ar[1] + ar2$ar[2] + ar2$ar[3]
stick_liking <- ar3$ar[1] + ar3$ar[2] + ar3$ar[3]
```

To summarise, stickiness measures for awareness, consideration and liking metrics are 0.764, 0.774 and 0.765, respectively.

### Another perspective:

In the brand-tracking / marketing-science sense, **stickiness** means the **staying power** of a metric like awareness, consideration, or liking. In plain English: after that metric goes up, **how much of the gain lasts** instead of quickly fading back toward normal? Hanssens et al. [@hanssens2014] define it as the inertia or lock-in of a change in the attitudinal metric. In their data, the funnel measures were advertising awareness, inclusion in the consideration set, and brand liking.

It is usually **estimated from time-series brand-tracking data**, not pulled from a single survey snapshot. The model is of the form\
$$
A_t = c \times A_{t-1}^{s} \times X_{1t}^{\beta_1} \times X_{2t}^{\beta_2} \times X_{3t}^{\beta_3} \times e^{u_t} ​,
$$

where $A_t$​ is the current awareness/consideration/liking measure, the X’s are marketing inputs, and $s$ is the **stickiness (carryover) parameter**. 

So, in short, **stickiness is the coefficient on the lagged attitude metric**. The paper says that if $s=0.60$, then about **60% of the change carries into the next period**; values closer to 1 mean a more persistent effect.

::: callout-important
So, boiled down:

**Stickiness = persistence of awareness/consideration/liking over time = the lagged carryover coefficient in the response model.** 
:::

## C. Responsiveness

Responsiveness, or **lift**, is the **short-term response of the attitude metric with respect to a marketing stimulus**.

-   We recommend using well-established, robust response functions to estimate responsiveness.

-   For example, the standard multiplicative response model produces elasticities as the following **responsiveness** metrics:

$$ 
Y_t = cY_{t-1}^\gamma X_{1t}^{\beta_1} X_{2t}^{\beta_2} e^{u_t}
$$

where ***Y*** is an attitude metric and $X_i$ (i=1,2) are marketing instruments. Not only do such response models provide readily interpretable results, but they have also been shown to outperform more complex specifications in forecasting product trials for consumer-packaged goods.

Note that *responsiveness* may be related to **potential** as follows: the closer the attitude metric is to its ceiling value, the more difficult it will be to register further increases through marketing.

-   To reflect that in the calculation, once we get the responsiveness of each mindset metric and sales, we can multiply it with the corresponding potential:

$$
\theta_i = \beta_i * Potential_i
$$

To estimate the responsiveness model, which is multiplicative, an easy way is to take the logarithm of both sides of the equation and estimate the model as a **log-linear model**. To be specific, we first take the log such that:

$$ 
log(Y_t) = c'+\gamma log(Y_{t-1}) + {\beta_1} log(X_{1t}) +  {\beta_2} log(X_{2t}) + {u_t} 
$$

Now, we examine the *log-linear model* for each mindset metric and sales to get responsiveness:

```{r warning=FALSE}
#| label: responsiveness-model
# We start by generating lagged variables for each mindset metric and sales

data_shampoo$lag_aware <- lag(data_shampoo$awareness)
data_shampoo$lag_aware[1] <- 0
data_shampoo$lag_liking <- lag(data_shampoo$liking)
data_shampoo$lag_liking[1] <- 0
data_shampoo$lag_consideration <- lag(data_shampoo$consideration)
data_shampoo$lag_consideration[1] <- 0
data_shampoo$lag_sales <- lag(data_shampoo$vol)
data_shampoo$lag_sales[1] <- 0

# Now estimate the log-linear model

response_aware <- lm(
  log(data_shampoo$awareness + 1) ~ log(lag_aware + 1) +
    log(data_shampoo$price + 1) +
    log(data_shampoo$promotion + 1) +
    log(data_shampoo$advertising + 1),
  data = data_shampoo
)

response_consideration <- lm(
  log(data_shampoo$consideration + 1) ~ log(lag_consideration + 1) +
    log(data_shampoo$price + 1) +
    log(data_shampoo$promotion + 1) +
    log(data_shampoo$advertising + 1),
  data = data_shampoo
)

response_liking <- lm(
  log(data_shampoo$liking + 1) ~ log(lag_liking + 1) +
    log(data_shampoo$price + 1) +
    log(data_shampoo$promotion + 1) +
    log(data_shampoo$advertising + 1),
  data = data_shampoo
)

response_sales <- lm(
  log(data_shampoo$vol + 1) ~ log(lag_sales + 1) +
    log(data_shampoo$price + 1) +
    log(data_shampoo$promotion + 1) +
    log(data_shampoo$advertising + 1),
  data = data_shampoo
)

#Summarize all the regression results here:

summary(response_aware)
summary(response_consideration)
summary(response_liking)
summary(response_sales)
```

Then, we **summarize** our estimation results in the below table:

```{r echo=FALSE}
#| label: responsiveness-summary-table
table_response_aware <- data.frame(round(response_aware$coefficients, 4))
table_response_liking <- data.frame(round(response_liking$coefficients, 4))
table_response_consideration <- data.frame(round(
  response_consideration$coefficients,
  4
))
table_response_sales <- data.frame(round(response_sales$coefficients, 4))

table_response_aware
table_response_consideration
table_response_liking
table_response_sales
```

## D. Conversion

-   An attitudinal metric would be meaningless for a brand if it does not impact sales. That is why we need to evaluate its power in terms of changing sales.

-   **Conversion** is the degree to which movements in the *attitudinal metric* convert to sales, similar to a conversion rate of leads into customer orders in B2B.

-   To calculate conversion, we again allow for a multiplicative funnel model that can be applied across different conditions.

For example, with intermediate attitudinal metrics awareness ($A_t$), consideration ($C_t$ ) and liking ($L_t$), a multiplicative funnel model for sales revenue ($S_t$) would be:

$$
S_t = cS_{t-1}^\gamma A_{t}^{\beta_1} C_{t}^{\beta_2} L_{t}^{\beta_3} e^{v_t}
$$

To estimate the conversion model, we again take the log of both sides of the equation above and estimate a linear model.

```{r}
#| label: conversion-model
conversion <- lm(
  log(data_shampoo$vol + 1) ~ log(lag_sales + 1) +
    log(data_shampoo$awareness) +
    log(data_shampoo$consideration) +
    log(data_shampoo$liking),
  data = data_shampoo
)
summary(conversion)
```

Summarizing model results, we get:

```{r echo=FALSE}
#| label: conversion-table
table_conversion <- data.frame(
  round(conversion$coefficients, 4)
)

table_conversion
```

## Summary results

Now, we are finally done with all the steps for analysis. We can aggregate the results from each stage of the estimation in the following table so that it is easier to see the measures on potential, stickiness, responsiveness and conversion.

```{r echo=FALSE}
#| label: final-table
table_final <- data.frame(
  Item = c(
    "beginning level",
    "potential",
    "stickiness",
    "responsiveness to advertising",
    "responsiveness to promotion",
    "conversion"
  ),
  awareness = c(
    round(mean(data_shampoo$awareness) / 100, 3),
    round(1 - mean(data_shampoo$awareness) / 100, 3),
    round(stick_awareness, 3),
    round(response_aware$coefficients[5], 3),
    round(response_aware$coefficients[4], 3),
    round(conversion$coefficients[3], 3)
  ),
  consideration = c(
    round(mean(data_shampoo$consideration) / 100, 3),
    round(1 - mean(data_shampoo$consideration) / 100, 3),
    round(stick_consideration, 3),
    round(response_consideration$coefficients[5], 3),
    round(response_consideration$coefficients[4], 3),
    round(conversion$coefficients[4], 3)
  ),
  liking = c(
    round(mean(data_shampoo$liking_transformed) / 100, 3),
    round(1 - mean(data_shampoo$liking_transformed), 3),
    round(stick_liking, 3),
    round(response_liking$coefficients[5], 3),
    round(response_liking$coefficients[4], 3),
    round(conversion$coefficients[5], 3)
  )
)

table_final
```

## Appeal

Finally, the appeal of each marketing mix action (i.e. promotion and advertising) is calculated by multiplying the potential, responsiveness, long-run multiplier that uses stickiness, and conversion.

$$
Appeal = Potential * Reponsiveness_{short-run} * 1/(1-Stickiness) * Conversion
$$

```{r echo=FALSE}
#| label: appeal-table
#calculate appeal from model estimations; Note that we do not multiply potential here since responsiveness we get from the model already incorporates potential.

# Appeal of advertising on awareness
appeal_adv_awareness <- round(
  table_final[2, "awareness"] *
    table_final[4, "awareness"] *
    (1 / (1 - table_final[3, "awareness"])) *
    table_final[6, "awareness"],
  3
)

# Appeal of promotion on awareness
appeal_promo_awareness <- round(
  table_final[2, "awareness"] *
    table_final[5, "awareness"] *
    (1 / (1 - table_final[3, "awareness"])) *
    table_final[6, "awareness"],
  3
)

# Appeal of advertising on consideration
appeal_adv_consideration <- round(
  table_final[2, "consideration"] *
    table_final[4, "consideration"] *
    (1 / (1 - table_final[3, "consideration"])) *
    table_final[6, "consideration"],
  3
)

# Appeal of promotion on consideration
appeal_promo_consideration <- round(
  table_final[2, "consideration"] *
    table_final[5, "consideration"] *
    (1 / (1 - table_final[3, "consideration"])) *
    table_final[6, "consideration"],
  3
)

# Appeal of advertising on liking
appeal_adv_liking <- round(
  table_final[2, "liking"] *
    table_final[4, "liking"] *
    (1 / (1 - table_final[3, "liking"])) *
    table_final[6, "liking"],
  3
)

# Appeal of promotion on liking
appeal_promo_liking <- round(
  table_final[2, "liking"] *
    table_final[5, "liking"] *
    (1 / (1 - table_final[3, "liking"])) *
    table_final[6, "liking"],
  3
)

appeal_table <- data.frame(
  Item = c("appeal_advertising", "appeal_promotion"),
  awareness = c(appeal_adv_awareness, appeal_promo_awareness),
  consideration = c(appeal_adv_consideration, appeal_promo_consideration),
  liking = c(appeal_adv_liking, appeal_promo_liking)
)

appeal_table
```

::: callout-note
**Question 2**: What is your takeaway from these results? Which marketing tool (i.e. promotion or advertising) is more effective for the marketing manager of Zanten to improve consumer awareness, consideration and liking? If you were the manager of this shampoo brand, what type of marketing strategy would you like to adopt in general?
:::

# What-if analysis

Given that this chapter focuses on the value of marketing activities through consumer mindset metrics on brand sales, we would like to evaluate the impact of change in a particular marketing tool on long-term sales and decompose such effect into gain/loss due to mindset metrics and gain/loss due to transactions.

In particular, we will evaluate **two alternative options** for Zhang, the marketing manager of Zanten:

1.  Quintupling advertising spending

2.  Doubling promotion

## Quintupling advertising spending

::: {.callout-caution title="Scenario"}
Suppose that Zhang decides to quintuple advertising spending from *£100K* to *£500K* while keeping the rest of the marketing mix decisions same.
:::

Let us visualise the calculations by using the below table:

```{r echo=FALSE}
#| label: table-whatif
table_whatif <- data.frame(
  Item = c(
    "Advertising",
    "Promotion",
    "Awareness",
    "Consideration",
    "Liking",
    "Sales"
  ),
  Start = c(
    100,
    100,
    round(mean(data_shampoo$awareness) / 100, 3),
    round(mean(data_shampoo$consideration) / 100, 3),
    round(mean(data_shampoo$liking_transformed), 3),
    round(mean(data_shampoo$vol), 3)
  ),
  New = c(500, 100, "?", "?", "?", "?"),
  Gain = c(400, 0, "?", "?", "?", "?"),
  LRGain = c("/", "/", "?", "?", "?", "?"),
  Contribution = c("/", "/", "?", "?", "?", "/")
)


table_whatif
```

> The cells with *‘?’* are the ones that we need to fill in.
>
> To briefly summarise, we need to do the following:

1.  First, derive the new level of each mindset metric and sales based on the changes that we decide to make to the marketing mix (i.e. advertising).

2.  Then, by comparing the new and the old level, we can derive immediate or short-run gain for each metric and sale.

3.  Next, with the help of carryover parameters from responsiveness models, we transform those short-run gains to corresponding long-run gains.

4.  And lastly, to decompose the long-run gain in sales and attribute it to gains due to mindset metrics and gains due to transaction, we calculate how much those long-run gains from step 3 translate to an actual increase in sales.

### New values

From the table, the column *‘Start’* summarizes all the base values for each item. For example, the average level of awareness, expressed in percentage terms, is 27.5 per cent, leaving a 72.5 per cent potential.

Now, we need to decide what happens to all the mindset metrics and sales after we quintuple advertising (from 100K to 500K). First, we need to determine the new level of each metric and sale after the change to the marketing mix. More specifically, the change to marketing mix exerts its impact on each metric through short-run responsiveness. Therefore, we need to go back to our previous analysis and use the model parameters in responsiveness models. For mindset metric $M_i$ (i=1,2,3 in our case), we apply the formula below:

$$
New_{M_i} = Start_{M_i} *(Advertising_{new}/Advertising_{old})^{Responsiveness_{Advertising_{M_i}}})
$$

```{r warning=FALSE}
#| label: convert-to-numeric
# This code chunk is to make the variables numeric
table_whatif$Start <- as.numeric(table_whatif$Start)
table_whatif$New <- as.numeric(table_whatif$New)
table_whatif$Gain <- as.numeric(table_whatif$Gain)
table_whatif$LRGain <- as.numeric(table_whatif$LRGain)
table_whatif$Contribution <- as.numeric(table_whatif$Contribution)
```

```{r}
#| label: new-values
# Note that we do not include short run impact of change in promotion on each mindset metric because there is no change (100/100=1)

new_awareness <- round(
  (1 - table_potential[1, "value"]) *
    (table_whatif[1, "New"] / table_whatif[1, "Start"])^(table_final[
      4,
      "awareness"
    ]),
  3
)

new_consideration <- round(
  (1 - table_potential[2, "value"]) *
    (table_whatif[1, "New"] / table_whatif[1, "Start"])^table_final[
      4,
      "consideration"
    ],
  3
)

new_liking <- round(
  (1 - table_potential[3, "value"]) *
    (table_whatif[1, "New"] / table_whatif[1, "Start"])^table_final[
      4,
      "liking"
    ],
  3
)

#the responsiveness of sales to advertising can be found in the "response_sales"
new_sales <- round(
  mean(data_shampoo$vol) *
    (table_whatif[1, "New"] /
      table_whatif[1, "Start"])^as.numeric(response_sales$coefficients[5]),
  3
)
```

### Short-run gain

We derive the gain (loss if negative) of each mindset metric and sales by simply calculating the growth rate.

```{r}
#| label: short-run-gain
# Calculate gain (from "start" to "new")
gain_awareness <- round(
  (new_awareness / (1 - table_potential[1, "value"])) - 1,
  3
)
gain_consideration <- round(
  (new_consideration / (1 - table_potential[2, "value"])) - 1,
  3
)
gain_liking <- round((new_liking / (1 - table_potential[3, "value"])) - 1, 3)
gain_sales <- round((new_sales / mean(data_shampoo$vol)) - 1, 3)
```

### Long-run gain

Now, we can translate the short-run gain to long-run gain for each metric and sales performance.

-   To do this, we need to make use of the **carryover (gamma)** we get from responsiveness models (you may want to go back and check the tables named `table_responsiveness`). Again, for each mindset metric $M_i$ and sales:

$$
LRGain_{M_i} = Gain_{M_i}/(1-Carryover_{M_i})
$$

```{r}
#Calculate long run gain for each mindset metric and sales

lrgain_awareness <- round(
  gain_awareness / (1 - as.numeric(response_aware$coefficients[2])),
  3
)
lrgain_consideration <- round(
  gain_consideration / (1 - as.numeric(response_consideration$coefficients[2])),
  3
)
lrgain_liking <- round(
  gain_liking / (1 - as.numeric(response_liking$coefficients[2])),
  3
)
lrgain_sales <- round(
  gain_sales / (1 - as.numeric(response_sales$coefficients[2])),
  3
)
```

### Contribution

In the final step, we need to go back and refer to the conversion model (you may want to go back and check the table named ‘table_conversion’):

$$
Contribution_{M_i} = LRGain_{M_i} * Conversion_{M_i}
$$

```{r}
#| label: contribution
# Calculate conversion, i.e., decompose the gain and attribute to awareness, consideration, and liking

contribution_awareness <- round(
  lrgain_awareness * table_final[6, "awareness"],
  3
)
contribution_consideration <- round(
  lrgain_consideration * table_final[6, "consideration"],
  3
)
contribution_liking <- round(lrgain_liking * table_final[6, "liking"], 3)
```

### Final Output

Finally, we are able to fill in all the blank cells:

```{r echo=FALSE}
#| label: final-whatif-1
table_whatif <- data.frame(
  Item = c(
    "Advertising",
    "Promotion",
    "Awareness",
    "Consideration",
    "Liking",
    "Sales"
  ),
  Start = round(
    c(
      100,
      100,
      mean(data_shampoo$awareness) / 100,
      mean(data_shampoo$consideration) / 100,
      mean(data_shampoo$liking_transformed),
      mean(data_shampoo$vol)
    ),
    3
  ),
  New = round(
    c(500, 100, new_awareness, new_consideration, new_liking, new_sales),
    3
  ),
  Gain = round(
    c(400, 0, gain_awareness, gain_consideration, gain_liking, gain_sales),
    3
  ),
  LRGain = c(
    "/",
    "/",
    lrgain_awareness,
    round(lrgain_consideration, 3),
    round(lrgain_liking, 3),
    round(lrgain_sales, 3)
  ),
  contribution = c(
    "/",
    "/",
    round(contribution_awareness, 3),
    round(contribution_consideration, 3),
    round(contribution_liking, 3),
    "/"
  )
)

table_whatif
```

::: callout-tip
### Interpretation

By summing up $Contribution_{awareness}$, $Contribution_{consideration}$, and $Contribution_{liking}$, we can get the total long run gain in sales that originates from improvement in mindset metrics.

-   We have: 0.074 + 0.014 + 0.008 = 0.096

> We observe that:

-   Long term sales gain = 10.8%

-   Sales gain due to mindset metrics = 9.6%

-   Sales gain due to transactions = 10.8% - 9.6% = 1.2%

Therefore, we can tell that by quintupling advertising, there is a significant improvement in mindset metrics, and such improvement can contribute to a 9.6 per cent increase in sales. The eventual net effect on long-run sales is a 10.8 per cent increase, and this is due to a 1.2 per cent increase in sales due to transactions.
:::

Next, we will examine the scenario in which the brand manager of Zanten, Zhang, doubles the promotion spending.

## Doubling promotion spending

::: {.callout-caution title="Scenario"}
Suppose that the marketing manager of Zanten doubles the promotion spending from *£100K* to *£200K* and does not make further changes in the marketing mix. What would be the impact on long-term sales?
:::

Let us visualise the calculations that we need to make by using the table below:

```{r echo=FALSE}
#| label: table-whatif-2
table_whatif2 <- data.frame(
  Item = c(
    "Advertising",
    "Promotion",
    "Awareness",
    "Consideration",
    "Liking",
    "Sales"
  ),
  Start = c(
    100,
    100,
    round(mean(data_shampoo$awareness) / 100, 3),
    round(mean(data_shampoo$consideration) / 100, 3),
    round(mean(data_shampoo$liking / 7), 3),
    round(mean(data_shampoo$vol), 3)
  ),
  New = c(100, 200, "?", "?", "?", "?"),
  Gain = c(0, 100, "?", "?", "?", "?"),
  LRGain = c("/", "/", "?", "?", "?", "?"),
  Contribution = c("/", "/", "?", "?", "?", "/")
)

table_whatif2
```

> The cells with *'?'* are the ones that we need to fill in.

Briefly, we need to:

1.  Calculate the **new level of each mindset metric** and sales based on the changes that we decide to make to the marketing mix, i.e. doubling promotion spending.

2.  Then, by comparing the new and the old levels we will calculate the **short run gain** for each mindset metric and sales.

3.  Next, with the help of autoregressive (i.e. carryover) parameters from responsiveness models, we will transform those short-run gains to corresponding **long-run gains**.

4.  Finally, we will **decompose** the long-run sales gain into transaction route and mindset route.

### New Values

The column *'Start'* from the above table summarizes the base values for each metric. For example, the average level of awareness, expressed in percentage terms, is 21.51%, leaving a 78.49% potential.

Next, we will see what happens to all the mindset metrics and sales after we double promotion spending. To this end, we will determine the new level of each metric and sales after the change to the marketing mix variables (i.e.promotion). We need to use the estimated parameters from the responsiveness models. For mindset metric $M_i$ (i=1,2,3 in our case), we apply the formula below:

$$
New_{M_i} = Start_{M_i} *(Promotion_{new}/Promotion_{old})^{Responsiveness_{Promo_{M_i}}})
$$

```{r warning=FALSE}
#| label: convert-to-numeric-2
### This code chunk is to make the variables in the 'what-if' table numeric
table_whatif2$Start <- as.numeric(table_whatif2$Start)
table_whatif2$New <- as.numeric(table_whatif2$New)
table_whatif2$Gain <- as.numeric(table_whatif2$Gain)
table_whatif2$LRGain <- as.numeric(table_whatif2$LRGain)
table_whatif2$Contribution <- as.numeric(table_whatif2$Contribution)
```

```{r}
#| label: new-values-2
# Note that we do not include short run impact of change in ads on each mindset metric and sales because there is no change (100/100=1)

new_awareness2 <- round(
  (1 - table_potential[1, "value"]) *
    (table_whatif2[2, "New"] / table_whatif2[2, "Start"])^(table_final[
      5,
      "awareness"
    ]),
  3
)

new_consideration2 <- round(
  (1 - table_potential[2, "value"]) *
    (table_whatif2[2, "New"] / table_whatif2[2, "Start"])^table_final[
      5,
      "consideration"
    ],
  3
)

new_liking2 <- round(
  (1 - table_potential[3, "value"]) *
    (table_whatif2[2, "New"] / table_whatif2[2, "Start"])^table_final[
      5,
      "liking"
    ],
  3
)

#the responsiveness of promotion to sales can be found in the "response_sales"
new_sales2 <- round(
  mean(data_shampoo$vol) *
    (table_whatif2[2, "New"] /
      table_whatif2[2, "Start"])^as.numeric(response_sales$coefficients[4]),
  3
)
```

### Short-run gain

Next, we derive the gain (loss if negative) of each mindset metric and sales by simply calculating the *growth rate*.

```{r}
#| label: short-run-gain-2
#Calculate gain (from "start" to "new")
gain_awareness2 <- round(
  (new_awareness2 / (1 - table_potential[1, "value"])) - 1,
  3
)
gain_consideration2 <- round(
  (new_consideration2 / (1 - table_potential[2, "value"])) - 1,
  3
)
gain_liking2 <- round(
  (new_liking2 / (1 - table_potential[3, "value"])) - 1,
  3
)
gain_sales2 <- round(
  (new_sales2 / mean(data_shampoo$vol)) - 1,
  3
)
```

### Long-run gain

We can translate the short run gain into long run gain for each mindset metric and sales. To do this, we need to make use of the **stickiness (carryover) parameters** we get from the responsiveness models. You may want to go back and check the table named *'table_responsiveness'*. So, for each metric ($M_i$), we calculate the long-run gain as follows:

$$
LRGain_{M_i} = Gain_{M_i}/(1-Carryover_{M_i})
$$

```{r}
#| label: long-run-gain-2
#Calculate long run gain for each mindset metric and sales

lrgain_awareness2 <- round(
  gain_awareness2 / (1 - as.numeric(response_aware$coefficients[2])),
  3
)
lrgain_consideration2 <- round(
  gain_consideration2 /
    (1 - as.numeric(response_consideration$coefficients[2])),
  3
)
lrgain_liking2 <- round(
  gain_liking2 / (1 - as.numeric(response_liking$coefficients[2])),
  3
)
lrgain_sales2 <- round(
  gain_sales2 / (1 - as.numeric(response_sales$coefficients[2])),
  3
)
```

### Contribution

In the final step, we refer to the conversion model. You may want to go back and check the table named *'table_conversion'*. To find the contribution, we multiply the long run gain by the conversion parameter:

$$
Contribution_{M_i} = LRGain_{M_i} * Conversion_{M_i}
$$

```{r}
#| label: contribution-2
#Calculate conversion, i.e., decompose the gain and attribute to awareness, consideration, and liking

contribution_awareness2 <- round(
  lrgain_awareness2 * table_final[6, "awareness"],
  3
)
contribution_consideration2 <- round(
  lrgain_consideration2 * table_final[6, "consideration"],
  3
)
contribution_liking2 <- round(
  lrgain_liking2 * table_final[6, "liking"],
  3
)
```

### Final output

Finally, we are able to fill in all the blank cells:

```{r echo=FALSE}
#| label: final-whatif-table-2
table_whatif2 <- data.frame(
  Item = c(
    "Advertising",
    "Promotion",
    "Awareness",
    "Consideration",
    "Liking",
    "Sales"
  ),
  Start = round(
    c(
      100,
      100,
      mean(data_shampoo$awareness) / 100,
      mean(data_shampoo$consideration) / 100,
      mean(data_shampoo$liking / 7),
      mean(data_shampoo$vol)
    ),
    3
  ),
  New = round(
    c(100, 200, new_awareness2, new_consideration2, new_liking2, new_sales2),
    3
  ),
  Gain = round(
    c(0, 100, gain_awareness2, gain_consideration2, gain_liking2, gain_sales2),
    3
  ),
  LRGain = c(
    "/",
    "/",
    lrgain_awareness2,
    round(lrgain_consideration2, 3),
    round(lrgain_liking2, 3),
    round(lrgain_sales2, 3)
  ),
  contribution = c(
    "/",
    "/",
    round(contribution_awareness2, 3),
    round(contribution_consideration2, 3),
    round(contribution_liking2, 3),
    "/"
  )
)

table_whatif2
```

::: callout-tip
### Interpretation

By summing up $Contribution_{awareness}$, $Contribution_{consideration}$, and $Contribution_{liking}$, we can get the total long run gain in sales that originates from improvement in mindset metrics.

-   We have: 0.014 + 0.022 + 0.039 = 0.075

Given that there is a 24.5% gain in sales, we can tell that:

-   Long term sales gain = 24.5%
-   Sales gain due to mindset metrics = 7.5%
-   Sales gain due to transactions = 24.5% - 7.5% = 17%

> Recall that quintupling advertising expense resulted in a 10.8% increase in brand sales. The above analysis shows that doubling promotion results in a 24.5% increase in sales.

> Therefore, we can conclude that doubling promotion seems to be more effective in driving up sales.

> However, doubling promotion generates less improvement in consumer mindset metrics. If the ultimate objective of the marketing manager is to 'build a brand', then gains in mindset metrics should be more important. Therefore, depending on the strategic objective of the brand, Zhang, the marketing manager of Zanten, may want to choose either of these two strategies.

> To summarise, marketing managers can conduct such scenario analyses, compare the estimated long-run effects, and make evidence-based decisions to win their customers' minds, hearts, and sales.
:::

::: callout-note
**Question 3**: As Ms Zhang has to decide on the choices on two options, which of the two options should she propose to the CEO? Which marketing tool (i.e. promotion or advertising) is more effective for the marketing manager of Zanten to improve consumer awareness, consideration and liking? how should Ms Zhang justify her proposal to the CEO?
:::

# Saving and Loading Files

```{r}
#| eval: false
#| label: save-load-files
save.image("my_session.RData")
load("my_session.RData")
```

# References